Predicting Political Support in the United States

### 2024-03-12 output/paper.qmd
introduction complete

@@

```{r}
ces2022_1 |>
  ggplot(aes(x = education, fill = voted_for)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(gender)) +
  theme_minimal() +
  labs(
    x = "Highest education",
    y = "Number of respondents",
    fill = "Voted for"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

@@

In particular the models we had created are logistic regression, all prediciting who the participant voted for during the 2020 US Presidential Election.

@@

Finally, the last model aims to forecast a respondents vote knowing their prominent political ideals including education, gun ownership status, and student loan status. These are common political standpoints in the U.S.. Gun policy and student loan forgiveness is often a key political issue that separates the candidates. For the 2020 election, the candidates we are predicting for, Joe Biden and Donald Trump had opposing positions on student loan forgiveness. Whilst Trump was opposed to student loan forgiveness, backing incone-based repayment plans instead [@citetrumploanstance], Biden proposed to forgive a large amount of undergraduate tuition-related student debt [@citeballotpedia]. They also opposed each other on gun laws, with Trump advocating for the right to bear arms and Biden advocating gun safety policies aimed at holding manufacturers accountable [@citeballotpedia].

>>>>

<!-- The introduction is self-contained and tells a reader everything they need to know including:

1) broader context to motivate;

2) some detail about what the paper is about;

3) a clear gap that needs to be filled;

4) what was done;

5) what was found;

6) why it is important;

7) the structure of the paper.

A reader should be able to read only the introduction and know what was done, why, and what was found. Likely 3 or 4 paragraphs, or 10 per cent of total. -->

Voting for the leader of your country is one of the biggest decisions an American citizen can make. It decides many of the decisions that will be made throughout that four year term as well as how their life will change in that period of time. There may be changes that this elected government makes that they believe to be a great decision, but it may also be seen as terrible one in the eyes of others. This is why it is important to understand there are many factors that influence a voter's decision. They vote for the candidate who they believe will make the best decisions for them and their country. In this paper, we hope to show the impact of various factors on a voter's decision, and how these factors can be used to predict a voter's alignment. From personal stances and religious beliefs to political stances, or to biological factors, we hope to show how these factors can be used to predict a voter's alignment.

The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump (Gambino 2020). Using the 2022 Cooperative Election Study (CES) survey data, we aim to predict who a voter supported in the 2020 election based on their cultural background, political stances, and other factors. We will be using a representative sample of 60,000 Americans to discern patterns in these voters' responses  and how their personal backgrounds may have influenced their voting decision. We will be examining the impact of various factors such as: gender, race, educational background, religon, and personal stances on key political issues such as gun control and student loans. Our main focus is to compare the correlations in the 2022 CES survey data to the 2020 CES survey data, and to see if there are any significant correlations that can be used to predict a voter's alignment. Using these understandings, we hope to highlight the potential implications these correlations may have for this upcoming 2024 election.

With this context and information in mind, we developed this paper to be structured as follows. We will first present the data and the models we used to analyze the data. We will then present the results of our analysis, and discuss the implications of these results. Finally, we will conclude with a summary of our findings and the potential implications of these findings for the upcoming 2024 election. To go in more detail, after downloading and cleaning the 2022 CES survey data, we used the `rstanarm` [@citerstanarm] package to create a generalized linear model to predict who a voter supported in the 2020 election based on their cultural background, political stances, and other factors, as listed above. We then used the `modelsummary` [@citemodelsummary] package to summarize the results of our model. We found that there were strong correlations, with a strong example of this being the voter's personal stances on student loan forgiveness and gun ownership. With the two main parties, Democrats and Republicans, having opposing stances on these issues, we found that there was a strong correlation between these stances and who a voter supported. As Biden was in favor of student loan forgiveness, but opposed to gun ownership, while Trump was opposed to student loan forgiveness, but in favor of gun ownership, we discovered a correlation between a voter's ties to these stances because they directly oppose each other. This is just one example of the many correlations we found, and we will discuss these in more detail in the results section. These findings are important because it shows how a leading presidential party's stance on a certain topic has the potential to sway a voter's decision in the event that they align with that stance. This is important because it shows how a leading presidential party's stance on a certain topic has the chance to adjust and adapt to certain factors to possibly sway some more votes in their favour. We will also discuss the implications of these findings, and how they may be used to predict a voter's alignment in the upcoming 2024 election.

TODO: cite https://www.theguardian.com/us-news/2020/nov/07/joe-biden-wins-us-election-donald-trump-loses-final-result-2020

@@

```{r}
ces2022_1 |>
  ggplot(aes(x = education, fill = voted_for)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(gender)) +
  theme_minimal() +
  labs(
    x = "Highest education",
    y = "Number of respondents",
    fill = "Voted for"
  ) +
  coord_flip() +
  scale_fill_manual(values = c("Biden" = "blue", "Trump" = "red")) +
  theme(legend.position = "bottom")

@@

In particular the models we had created are logistic regression, all predicting who the participant voted for during the 2020 US Presidential Election.

@@

Finally, the last model aims to forecast a respondents vote knowing their prominent political ideals including education, gun ownership status, and student loan status. These are common political standpoints in the U.S.. Gun policy and student loan forgiveness is often a key political issue that separates the candidates. For the 2020 election, the candidates we are predicting for, Joe Biden and Donald Trump had opposing positions on student loan forgiveness. Whilst Trump was opposed to student loan forgiveness, backing incone-based repayment plans instead ()[@citetrumploanstance], Biden proposed to forgive a large amount of undergraduate tuition-related student debt ()[@citeballotpedia]. They also opposed each other on gun laws, with Trump advocating for the right to bear arms and Biden advocating gun safety policies aimed at holding manufacturers accountable ()[@citeballotpedia].
